---
title: "dplyr package"
output: html_notebook
---
```{r}
install.packages("dplyr")
library(dplyr)
```

```{r}
getwd()
mydata=read.csv("sampledata.csv")
```
## Selecting Random N Rows
```{r}
sample_n(mydata,3)
```
## Selecting Random Fraction of Rows
```{r}
sample_frac(mydata,0.1)
```
## Remove Duplicate Rows based on all the variables (Complete Row)

The distinct function is used to eliminate duplicates.
```{r}
x1=distinct(mydata)
```
## Remove Duplicate Rows based on a variable

The .keep_all function is used to retain all other variables in the output data frame.
```{r}
x2 = distinct(mydata, Index, .keep_all= TRUE)
```
## Remove Duplicates Rows based on multiple variables

we are using two variables - Index, Y2010 to determine uniqueness.
```{r}
x2 = distinct(mydata, Index, Y2010, .keep_all= TRUE)
```
## **select( ) Function**
Selecting Variables (or Columns)
```{r}
mydata2 = select(mydata, Index, State:Y2008)
```
 Dropping Variables

The minus sign before a variable tells R to drop the variable.
```{r}
mydata = select(mydata, -c(Index,State))
```
 Selecting or Dropping Variables starts with 'S' and 'Y'
The starts_with() function is used to select variables starts with an alphabet.
```{r}
mydata3=select(mydata,starts_with("S"))
mydata3plus=select(mydata,starts_with("Y"))
```
Adding a negative sign before starts_with() implies dropping the variables starts with 'Y' and 'I'
```{r}
mydata4=select(mydata,-starts_with("y"))
mydata4plus=select(mydata,-starts_with("I"))
```
ends_with()
```{r}
mydata5=select(mydata,ends_with("2008"))
```
Selecting Variables contain 'Y' in their names
```{r}
mydata5plus=select(mydata,contains("Y"))
```
Reorder Variables

The code below keeps variable 'State' in the front and the remaining variables follow that.
```{r}
mydata6=select(mydata,State,everything())
```
## **rename( ) Function**
Rename Variables
```{r}
mydata7=rename(mydata,Index1=Index)
```
##**filter( ) Function**
 Filter Rows

Suppose you need to subset data. You want to filter rows and retain only those values in which Index is equal to A.similarily, alaska.
```{r}
mydata8=filter(mydata,Index=="A")
mydata8plus=filter(mydata,State=="Alaska")
```
Multiple Selection Criteria

The %in% operator can be used to select multiple items.
```{r}
mydata9=filter(mydata,Index%in%c("A","C"))
```















